% What we did
In this experiment, we compute errors in estimated orientation for different features and regressors in order to determine the best combination for the retinogram dataset. To account for any bias introduced by an uneven distribution of thickness or contrast over the vessel set, we also investigate orientation error with respect to these image properties. This is important in this application since we are particularly interested in measuring the properties of the thinnest and faintest vessels.

In the following experiments, 200\,000 training pairs were taken from pixels that were known (from the ground truth segmentation) to lie on a vessel; the orientation of background pixels was not defined in this work.

To compare the performance of different features and regressors, we generated the cumulative distribution function over error (\fref{f:retina_graphs}a) and summarised performance by the median\footnote{The median is more robust than the mean to outliers in the error distribution.} errors over both the whole vessel and over the centre line only (\tref{t:retinopathy}).

We define ground truth orientation, at the vessel centres by skeletonising the masks and computing a locally linear fit; we then assign to every vessel pixel the orientation of its nearest centreline pixel.
\comment{Is this a reasonable proxy for true orientation?}

As a preprocessing step, we transformed all images to monochrome via a weighted sum of the three RGB channels (this proved marginally more successful than extracting features from all or any of the channels individually).

\comment{How can we justify throwing away two thirds of the available data? What were the weights? How much more successful was this? Is the difference significant?}

When sampling pixels with which to form the training data, we selected 200\,000 vessel pixels randomly over the set of training images. At every selected pixel, we then paired the chosen filter responses with the corresponding ground truth orientation, $t_{gt}$.

\comment{Why 200k points? Was this limit dictated by system requirements? What effect does the number of points have? (Try halving and doubling the number)}




\input{experiments/measuring_orientation/in_retinograms/fig_retinogram_orientation.tex} % from BMVC
\input{experiments/measuring_orientation/in_retinograms/fig_retinogram_wrt_properties.tex} % from MICCAI

% Discussions

\subsubsection{Choice of feature}
\input{experiments/measuring_orientation/in_retinograms/wrt_feature.tex}

\subsubsection{Choice of regressor}
\input{experiments/measuring_orientation/in_retinograms/wrt_regressor.tex}

\subsubsection{Effect of window size}
\input{experiments/measuring_orientation/in_retinograms/wrt_window_size.tex}

\subsubsection{Effects of line thickness}
\input{experiments/measuring_orientation/in_retinograms/wrt_width.tex}

\subsubsection{Effects of line contrast}
\input{experiments/measuring_orientation/in_retinograms/wrt_contrast.tex}

% We conclude that...?
Overall, combining \dtcwt{} features with Random Forest regression was most successful with the \dtcwt{} in particular benefiting from the improvement of Random Forests over simple linear regression (\tref{t:retinopathy})
